---
title: "Tweet Classifier"
author: "Matteo Larrode"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Report on the Sentiment About Airlines Using Text Data Analysis on Tweets

## Introduction
This report analyzed more that 11,000 tweets to gauge customers’ sentiment towards the company and its major competitors. Word usage is compared between positive and negative tweets, and across airlines. Then, two predictive tools are built to classify them. A dictionary-based method and one using supervised machine-learning are compared to help the company respond better to customers in real time.

```{r prep, include=FALSE}
#Prep Work & Packages
load("data/tweets.Rda")

library(quanteda)
library(quanteda.textstats)
library(quanteda.textplots)
library(tidyverse)
library(glmnet)
library(ggplot2)
```

## Text preparation

Text preparation started with the removal of numbers, stopwords, punctuation, and symbols commonly used on Twitter. Very rarely used words were also removed to ease text data analysis. Term Frequency–Inverse Document Frequency (TF-IDF) weighting was carried out to help give an accurate picture of words that are most characteristic of tweets, and diminish the importance of uninformative words. TF-IDF weightings were carried out after the different groupings of documents as a TF-IDF weighted document-term matrix should not be amalgamated.

```{r text prep, warning=FALSE}
#Text Preparation ----------
tweetCorpus <- corpus(tweets$text, docvars = tweets)

#turn corpus into document-term matrix; make everything lower-case, remove numbers, punctuation and stopwords; TF-IDF weighting; remove airline names
dfm_tweets0 <- tweetCorpus %>%
  tokens(remove_numbers=T,
         remove_punct=T,
         include_docvars=T,
         remove_symbols = T) %>%
  tokens_remove(stopwords("en")) %>%
  dfm(tolower=T)%>%
  dfm_tfidf()

#eliminate rare words & other words
dfm_tweets <- dfm_tweets0 %>% 
  dfm_trim(min_docfreq = 5) %>% 
  dfm_remove(c("@united", "@usairways", "@americanair", "@southwestair", "@jetblue", "@virginamerica", "flight", "amp", "aa", "get", "now", "united", "jetblue", "us"))

cat("Document-Term Matrix after pre-processing\n", 
    "Number of tweets:", dim(dfm_tweets)[1], 
    "\nNumber of words: ", dim(dfm_tweets)[2])
```

## Word Usage Analysis

### Sentiment
```{r message=FALSE, warning=FALSE}
# Get the frequency of terms in the document-term matrix
freq <- textstat_frequency(dfm_tweets, force = TRUE)
top_15_terms <- freq[1:15, 1]

cat("Top 15 most used words:\n", top_15_terms, sep = "\n")

tweets <- tweets %>%
  mutate(sentiment_long = ifelse(sentiment == 0, "Positive", "Negative"))

# Compute the table of sentiment frequencies for each airline
sentiment_table <- tapply(tweets$sentiment_long, tweets$airline, table)

# Loop through each airline and print the proportion of each sentiment
for (airline in names(sentiment_table)) {
  cat("Airline:", airline, "\n")
  sentiment_count <- sentiment_table[[airline]]
  sentiment_prop <- prop.table(sentiment_count)
  print(sentiment_prop)
  cat("\n")
}


# Words most associated with negative & positive sentiment
bysentiment <- textstat_frequency(dfm_tweets,25,groups=sentiment, force=T) %>%
  mutate(sentiment_long = ifelse(group == 0, "Positive", "Negative"))


#Wordclouds
pos_tweets <- dfm_tweets[dfm_tweets$sentiment == "0",]
textplot_wordcloud(pos_tweets,
                   min_size = 0.5,
                   max_size=6,
                   color = "darkblue",
                   comparison=FALSE, 
                   max_words=50)

neg_tweets <- dfm_tweets[dfm_tweets$sentiment == "1",]

textplot_wordcloud(neg_tweets,
                   min_size = 0.5,
                   max_size=6,
                   color = "red",
                   comparison=FALSE, 
                   max_words=50)


```
Some words are commonly associated with a positive or negative sentiment. Words of gratitude like “thanks” and “appreciate”, along with “awesome”, “great” and “amazing” are found in many positive tweets. On the other hand, the language associated to delay or cancellation (“hours”, “cancelled”, “hold”, “delayed”, “time”) are commonly associated with negative tweets.

### Sentiment by airline

While commonly used words associated with a positive sentiment are similar across airlines, an overview of how negative word usage differs across airlines can help reveal specific customer complaints. For example, American Airlines, Southwest, and US Airways customers seemed to have troubles with cancelled flights and help service. JetBlue and United customers mostly complain about delays. Virgin America customers seem to dislike the website of the company. Many negative tweets associated to United mention “bag”, hinting at problems of luggage management by the company.






